###### IODS - Introduction to Open Data Science - 2017

# Week 2:
##Regression and model validation
### Brief explanation of the data
The dataset analysed concerns the learning experience of a set of students. Students were asked questions about their approach to studying as well as their attitude towards it. In particular 3 kinds of approach were pinpointed: deep, strategic and surface approaches. The dataset also contains personal information about the participants: gender and age. The collected data were compared to exam points.  

The following command reads the data into R. As a matter of convenience, the dataset is loaded from a file on the internet, although the Data wrangling part has been completed (check file data/create_learning2014.txt and data/learning2014.txt in my github repository).
```{r}
learning2014 <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", sep=",", header=TRUE)
```
  
Let us take a look at the structure of the dataset: 
```{r}
str(learning2014)
```
The dimensions are: 166 observations and 7 variables. Those are: gender, age, attitude, deep (deep learning approach), stra (strategic learning approach), surf (surface learning approach) and points (exam grade) for each observed student. 
  
### Graphical overview of data and summary
Using the library 'ggplot2', let us examine the graphical overview of our data:
```{r}
library(ggplot2)
ggplot(learning2014, aes(x = attitude, y = points, col = gender)) + geom_point() + geom_smooth(method = "lm") + ggtitle("Students' attitude versus exam points")
```
  
The above plot displays students' attitudes on the x-axis and and their exam points on the y-axis and shows the correlation between the two. In addition, the different colours display the students' gender. Gender does not seem to play an important role with respect to exam points, as both female and male participants are equally distributed throughout the plot. On the other hand, the student's attitude towards learning is crucial for what concerns exam points, showing that in general the higher the attitude, the higher the exam points.
  
Let us take a look at summaries for variables in our dataset with the function summary():
```{r}
summary(learning2014)
```
  
Notice that female observations are roughly double than male's, but they are equally distributed throughout the graph, as we have seen in the previous plot. The age of paticipants is ranging between 17 and 55 years old, although the 3rd Quartile is 27.00 which indicates that the vast majority of observatons falls below the middle point in this range. Range of attitudes is 1.400 to 5.000, but the vast majoity of observations lies in the middle, between values 2 and 4. Similarly, range of exam points is 7.00 to 33.00, but the vast majority falls between values 16 and 30. This suggests that exam points may be dependent on the attitude.

We are now analysing three types of learning approaches, i.e. deep, strategic and surface learning. Deep-learning range is 1.583 to 4.917; strategic-learning range is 1.250 to 5.000; surface learning range is 1.583 to 4.333. From wider to narrower range the order is: strategic, then deep and surface last. Deep learning is the learning approach with higher summary values, followed by strategic and surface at last with the lowest values.

### Linear regression
Let us assume that exam points are dependent on 3 variables, for instance attitude, strategic learning and age:
```{r}
my_model <- lm(points ~ attitude + stra + deep, data = learning2014)
summary(my_model)
```

The residuals are fairly high for what concerns min and max values, but they get closer to 0 at the median. Std. Error falls between 0 and 1 for all 3 independent variables. T value for the 3 variables is highest for attitude and lowes for deep learning, while strategic learning lies in between. A multiple R-squared of 0.2097 indicates that the model fits the data for roughly 21% of the observations.

### Residual vs. Fitted values
```{r}
plot(my_model, 1)
```

The above plot illustrates the variation around the estimated regression line. Since this variation is constant, this suggests that the assumption of equal error variances is reasonable and that the assumption that the relationship is linear is also reasonable.

### Normal QQ-plot
```{r}
plot(my_model, 2)
```

The above plot illustrates the normal distribution of our observations. Despite a tail on the left and a skew on the right, the circles in the scatterplot lie all quite close to the line, suggesting that our assumpton is plausible.

### Residuals vs. Leverage
```{r}
plot(my_model, 5)
```

The above plot shows there are no observations that are outside Cook's distance. Since it is not the case for high Cook's distance, this means there are no influential cases to the regression results. This once more backs up our hypothesis for a model.

Therefore, we can conclude that our model of exam points dependent on attitude, deep and strategic learning must correspond to reality.